BIP Architectures and Open pNets

### 2019-01-25 2018-Arch-pNets/arch-pNets.tex

\usepackage{soul}
\soulregister\cite7
\soulregister\ref7
\soulregister\pageref7
\soulregister{\em}{0}
\soulregister{\mdash}{0}
\soulregister{\st}{7}
\soulregister{\bf}{0}
\soulregister{\fig}{7}
\soulregister{\ie}{0}
\soulregister{\etc}{0}
\soulregister{\eg}{0}
\soulregister{\wrt}{0}
\soulregister{\resp}{0}
\soulregister{\cf}{0}

@@

@@

\newcommand{\mdash}[1][]{---#1}
\newcommand{\ndash}{--}
\newcommand{\ie}[1][\ ]{i.e.#1}
\newcommand{\etc}[1][\ ]{etc.#1}
\newcommand{\eg}[1][\ ]{e.g.#1}
\newcommand{\cf}[1][\ ]{cf.#1}
\newcommand{\wrt}[1][\ ]{w.r.t.#1}
\newcommand{\resp}[1][\ ]{resp.#1}

@@

Finally, we assume the existence of a universal data domain
given as a partially-ordered set $(\data, \order)$, potentially
encompassing several copies of any given data type with
different orders.  For example, we assume that $(\data,
\order)$ comprises both the unordered set of Booleans $\sB =
\tuple{\{\true, \false\}, \emptyset}$ and the naturally ordered one
$\sB^< = \tuple{\{\true, \false\}, \{\false < \true\}}$, and
similarly for integer and real numbers; as well as the set of
intervals ordered by inclusion.

For a set of variables $V \subseteq \variables$, we denote
$\valuations{V} \bydef{=} \{\val{}{}: V \rightarrow \data\}$
the set of \emph{valuations} of the variables in $V$.
Valuations extend canonically to expressions.  Thus
$\val{}{}(x)$ and $\val{}{}(e)$ denote, respectively, the
values of the variable $x$ and the expression $e$ under the
valuation $\val{}{}$.
%
\addSB{
%
It is important to observe that in practice, and in the examples of
this paper, all the variables are sorted, meaning that their values
can only range over given subsets of $\data$, such as $\sB$ or
$\sB^<$.  Hence, applied to expressions, valuations should be
interpreted as partial functions $\exprs{V} \rightharpoonup
\data$\footnote{
%

%
}
%
}

For a valuation $\val{}{} \in
\valuations{V}$ and an assignment $(x_i := e_i)^{i \in I} \in
\assigns{V}$, we denote $\val{}{}[(x_i := e_i)^{i \in I}]$ the
valuation defined by
\[
\val{}{}\bigl[(x_i := e_i)^{i \in I}\bigr](x) \bydef{=}
\begin{cases}
  \val{}{}(x), & \text{if } x \not\in x_i^{i \in I}\,,\\
  \val{}{}(e_i), & \text{if } x = x_i, \text{for some } i \in I\,.
\end{cases}
\]
For two valuations $\val{1}{}, \val{2}{} : V \rightarrow
\data$, we denote \[\valdiff{\val{1}{}}{\val{2}{}} \bydef{=}
\bsetdef{v \in V}{\val{1}{}(v) \neq \val{2}{}(v)}\] the set of
variables that are assigned different values by the two
valuations.  As usual, we write $\val{1}{} \order \val{2}{}$
iff $\val{1}{}(x) \order \val{2}{}(x)$, for all $x \in V$.  An
expression $e \in \exprs{V}$ is called \emph{monotonic} if, for
any two valuations $\val{1}{}, \val{2}{} \in \valuations{V}$,
$\val{1}{} \order \val{2}{}$ implies $\val{1}{}(e) \order
\val{2}{}(e)$.
%
Similarly, an assignment $(x_i := e_i)^{i \in I} \in \assigns{V}$ is
called monotonic if all expressions $e_i^{i \in I}$ are monotonic.
We will denote $\guards{V} \subset \boolexprs{V}$, $\monexprs{V}
\subset \exprs{V}$ and $\updates{V} \subset \assigns{V}$ the sets of
monotonic Boolean and generic expressions and assignments,
respectively.
%

@@

\noteSBin{%
To guarantee preservation of properties guards and expressions
have to be \emph{monotonic}.  In the case of a trivial order, \ie
when nothing is comparable, \emph{any guard or expression is
  monotonic}.  Thus, this is not really a constraint!

Do we put this in the definition of expressions and guards or in
the definitions of components and interaction models?
}

\begin{definition}[Components]
  \label{defn:component}
  A \emph{component} is a tuple $\tuple{Q, q^0, V, \val{0}{}, P,
  \export, \goesto{}}$, where
  \begin{itemize}
  \item $Q$ is a set of \emph{states}, with $q^0 \in Q$ the
    \emph{initial state}, 
  \item $V$ is a set of \emph{component variables},
  \item $\val{0}{} : V \rightarrow \data$ is an \emph{initial
    valuation} of the component variables, 
  \item $P$ is a set of \emph{ports}, with
    $\export : P \rightarrow 2^V$ 
    the \emph{set of variables exported by each port},
  \item $\goesto{}\, \subseteq
    Q \times 2^P \times \guards{V} \times \updates{V} \times Q$
%
    is a \emph{transition relation}, with transitions
    labelled by \emph{interactions}, \ie triples consisting of a non-empty set of ports
    $\emptyset \neq a \subseteq P$, a monotonic Boolean \emph{guard} $g \in
    \guards{V}$ and a monotonic \emph{update assignment} \hl{$u \in
    \updates{V}$}.
  \end{itemize}
%
  We call the triple $(V,P, \export)$ the \emph{interface} of the
  component.\,\footnote{%
%
  In practice, only exported variables, \ie those in the set
  $\bigcup_{p \in P}\export(p)$, belong to the component interface and
  can be used for the interactions with other components (see
  \defn{im}).  However, in the context of this paper we can simplify
  by omitting this separation.
%
  }
%
  Notations $q \goesto{a, g, u} q'$ and $q \goesto{a, g, u}$ are as usual;
  for a component $B$, we denote $Q_B$, $q^0_B$, $V_B$, $\val{0}{B}$, $P_B$,
  and $\export[B]$ the corresponding constituents of $B$.  We will
  skip the index on the transition relations $\goesto{}$, since it
  is always clear from the context.
\end{definition}

\begin{figure}[t]
  \centering
  \includegraphics[width=\columnwidth]{BIPspec-ArchFailureTimerMax-v4}
  \caption{The BIP specification of the Failure Monitor architecture}
  \label{fig:schema:ArchFailure:BIP}
\end{figure}

@@

The model in \fig{schema:ArchFailure:BIP} has two components:
%% The
%% component $\NameOprnd$ is an \emph{operand} of the architecture, \ie
%% it only has its interface defined: $(\emptyset, \{\PortFinish,
%% \PortResume, \PortFail\}, \export[\emptyset])$, with
%% $\export[\emptyset](p) = \emptyset$, for all $p \in \{\PortFinish,
%% \PortResume, \PortFail\}$.  
%% The other two components,
%
$\NameTimer$ and $\NameCtrl$, with interfaces
\[\bigl(
\{t, z\},
\{\PortTick, \PortCancel, \PortTOT\},
\export[\NameTimer]
\bigr)\,,
\qquad
\export[\NameTimer](p) =
\begin{cases}
  \{t, z\}, & \text{if } p = \PortTick,\\
  \emptyset, & \text{otherwise}
\end{cases}
\]
and
\[\bigl(
\{\mathit{zone}\},
\{\PortReset, \PortFail, \PortResume, \PortTOC\},
\export[\NameCtrl]
\bigr)\,,
\quad
\export[\NameCtrl](p) =
\begin{cases}
  \{\mathit{zone}\}, & \text{if } p = \PortFail,\\
  \emptyset, & \text{otherwise}
\end{cases}
\]
%
respectively. %% , where $\export[\NameCtrl](\PortFail) =
%% \{\mathit{zone}\}$ and $\export[\NameCtrl](p) = \emptyset$, for all
%% other ports $p$; $\export[\NameTimer](\PortTick) = \{t, z\}$ and
%% $\export[\NameTimer](p) = \emptyset$, for all other ports $p$.
Variable $t$ is implicitly assumed to be an integer.  Variables $z$
and $\mathit{zone}$ are implicitly assumed to be integer intervals of
the form $[l,u]$.  For both types, we also implicitly assume the
natural ordering: increasing integers and interval inclusion.

%% are \emph{coordinators}\mdash they are specified
%% completely, including the states and transitions defining their
%% \emph{behaviour}.
Component \emph{behaviour} is defined by states ($\{\cstate{0},
\cstate{1}, \cstate{2}\}$, for $\NameCtrl$; $\{\tstate{0}\}$, for
$\NameTimer$) and transitions.
%
Transitions are labelled with the ports of the
corresponding components, Boolean guards and update assignments on local
variables.  For instance, the loop transition $\tstate{0}
\goesto{\PortTick, \GuardTick[], \TimerUpd[]} \tstate{0}$ in the
$\NameTimer$ component is labeled by the port $\PortTick$, it can be
fired only when the current value of the local variable $t$ is less
than $z.u$.
%% , where $z = [l, u]$ is a local variable representing an
%% interval
Upon firing, this transition increments the value of $t$ by $1$.  The
guards and update assignments of the transitions of $\NameCtrl$ are
omitted.  The default values are the constant predicate $\true$ (true) and
the empty assignment {\noop}, respectively.  The initial states
\tstate{0} and \cstate{0}, and valuations $\val{0}{\NameTimer} = \{t
\mapsto 0, z \mapsto \topinterval\}$ and $\val{0}{\NameCtrl} =
\{\mathit{zone} \mapsto [\mathrm{Min}, \mathrm{Max}]\}$ are shown by
incoming arrows $\goesto{\TimerInit[]} \tstate{0}$ and
$\goesto{\IntervalInit[]} \cstate{0}$, respectively.\footnote{%
%
  We use $\topinterval$ to denote the top interval in the inclusion
  order, \ie $(-\infty, +\infty)$.
%
} The constants $\mathrm{Min}$ and $\mathrm{Max}$ are the parameters
of the architecture.
%\noteLH{It looks like Max is in zone, I do not understand}

\todoSBin{Speak of guard monotonicity}

\begin{definition}[Component semantics]
  \label{defn:comp:semantics}
  The \emph{open semantics} of a component $B = \tuple{Q, q^0, V,
  \val{0}{}, P, \export, \goesto{}}$ is given by the LTS denoted $\semopen{B} = \tuple{S,
  s^0, \goesto{}}$, where $S = Q \times \valuations{V}$, $s^0 =
  \tuple{q^0, \val{0}{}}$ and $\goesto{}$ is the minimal transition
  relation satisfying the rule
  %
  \begin{equation}
    \label{eq:comp:semantics}
    \infer{
      q \goesto{a, g, u} q'
      \and
      \val{}{} \models g
      \and
      \val[\primeit]{}{} = \tilde{\val{}{}}[u]
      \and
      \valdiff{\val{}{}}{\tilde{\val{}{}}} \subseteq \export(a)
    }{
      (q, \val{}{}) \goesto{a,\tilde{\val{}{}}} (q', \val[\primeit]{}{})
    }
    \,.
  \end{equation}
  %
  %% \todoSB{Do we actually need this?}\todoLH{I agree: I do not think we need this}
  %% \hl{The \emph{closed semantics} of $B$ is given by the LTS denoted
  %% $\semclosed{B} \bydef{=}\reachable{\semopen{B}}$, comprising
  %% only the reachable states of $\semopen{B}$.}
\end{definition}

@@

\begin{definition}[Interaction model]
  \label{defn:im}
  Consider a finite set of component interfaces $(V_i, P_i,
  \export[i])^{i \in I}$, such that all $P_i$ and all $V_i$ are
  pairwise disjoint, \ie $\forall i \neq j,\ P_i \cap P_j = V_i \cap
  V_j = \emptyset$.  Let $P = \bigcup_{i \in I} P_i$, $V = \bigcup_{i
    \in I} V_i$ and $\export : P \rightarrow 2^V$ such that, for any
  $p \in P_i$, $\export(p) = \export[i](p)$. 
%
  %% \noteSB{Updated this}\hl{and $V_a \bydef{=} \bigcup_{i \in
  %%   \supp{a}} \export[i](a \cap P_i)$ the set of component
  %% variables exported through $a$.}

@@

\begin{definition}[Composition]
  \label{defn:im:sem}
  The \emph{composition} of a finite set of components $\cB = \tuple{Q_i,
  q^0_i, V_i, \val{0}{i}, P_i, \export[i], \goesto{}}^{i \in I}$ with
  the interaction model $\Gamma$ over $\tuple{V, P, \export}$ (as in \defn{im})
  is the component $\Gamma(\cB) = \tuple{Q, q^0, V, \val{0}{}, P,
    \export, \goesto{}}$, where
%
  $Q = \prod_{i \in I} Q_i$;
%
  $q^0 = (q_i^0)^{i \in I}$;
%
  $\val{0}{}: V \rightarrow \data$ is such that, for any $v \in
  V_i$, $\val{0}{}(v) = \val{0}{i}(v)$;
%
  and $\goesto{}$ is the minimal transition relation satisfying
  the rule
%
  \begin{gather}
%%     \label{eq:im:empty}
%%     \infer{
%%       q_j \goesto{\emptyset, g, e} q_j'
%%       \and
%%       \forall i \neq j, q_i = q_i'
%%     }{
%%       (q_i)^{i \in I} \goesto{\emptyset, g, e} (q_i')^{i \in I}
%%     }\,,
%% %
%%     \\[0.5\baselineskip]
    \label{eq:im:sem}
%
    \infer{    
      (a, g, u) \in \Gamma
      \and
      a \neq \emptyset
      \and
      \forall i \in \supp{a}, q_i \goesto{a \cap P_i, g_i, u_i} q_i'
      \and
      \forall i \not\in \supp{a}, q_i = q_i'
%
      \\\\
      \textstyle
%
      g' = g \land \bigwedge_{i \in \supp{a}} g_i
      \and
      u' = u; u_i^{i \in \supp{a}}
    }{
      (q_i)^{i \in I} \goesto{a, g', u'} (q_i')^{i \in I}
    }\,.
  \end{gather}
  \todoSBin{Add the intuitive explanation}
\end{definition}

@@

Connectors are hierarchical, tree-like structures with component ports
at the leaves.  They define sets of interactions, based on the
attributes of the connected ports, which may be either \emph{trigger}
(triangles in \fig{schema:ArchFailure:BIP}) or \emph{synchron}
(bullets in \fig{schema:ArchFailure:BIP}).
%
If all sub-connectors of a connector are synchrons, then an
interaction may be executed by the connector only if each subconnector
can contribute.
%
If at least one of the sub-connectors is a trigger, then any
interaction consisting of contributions of any set of sub-connectors,
\emph{involving at least one of the triggers}, can be executed.
The composition semantics of BIP systems consists in
firing exactly one interaction, enabled through at least one of the
top-level connectors, at each execution round.

\todoSB{Update this}
\hl{For instance, in \fig{schema:ArchFailure:BIP}, the two ports
$\NameTimer.\PortStart$ and $\NameCtrl.\PortFail$ are always
synchronised, since they belong to exactly one binary sub-connector,
where they are both synchrons.  In particular, this means that
whenever the transition \cstate{0} \goesto{\PortFail} \cstate{1} is
fired, so is the transition \tstate{0} \goesto{\PortStart,
  \TimerInit[]} \tstate{1}, initialising the timer.  The binary
connector $\NameTimer.\PortStart \twosynch
\NameCtrl.\PortFail$ is a sub-connector of a hierarchical connector,
where the port $\NameOprnd.\PortFail$ is a trigger.  Thus, the above
interaction can only happen together with $\NameOprnd.\PortFail$,
forming a ternary interaction.  On the contrary, being a trigger, the
port $\NameOprnd.\PortFail$ can fire alone, forming a singleton
interaction.}

\todoSBin{Speak of data transfer}

\todoSBin{Provide the interaction model}

@@

\begin{note}
  \label{rem:maxprogress}
  Below, we will assume that the maximal progress priority $\mu$ is
  applied after each composition with an interaction model.
\end{note}

@@

\todoSBin{Sort out the following two paragraphs}

>>>>

\usepackage{soul}
\soulregister\cite7
\soulregister\ref7
\soulregister\pageref7
\soulregister{\em}{0}
\soulregister{\mdash}{0}
\soulregister{\st}{7}
\soulregister{\bf}{0}
\soulregister{\fig}{7}
\soulregister{\ie}{0}
\soulregister{\etc}{0}
\soulregister{\eg}{0}
\soulregister{\Eg}{0}
\soulregister{\wrt}{0}
\soulregister{\resp}{0}
\soulregister{\cf}{0}

@@

\newcommand{\btuple}[2][\ensuremath{,\allowbreak}]{%
  \setTupleDelims[\bigl(]{\bigr)}%
  \tuplebase[#1]{#2}%
}

@@

\newcommand{\mdash}[1][]{---#1}
\newcommand{\ndash}{--}
\newcommand{\ie}[1][\ ]{i.e.#1}
\newcommand{\etc}[1][\ ]{etc.#1}
\newcommand{\eg}[1][\ ]{e.g.#1}
\newcommand{\Eg}[1][\ ]{E.g.#1}
\newcommand{\cf}[1][\ ]{cf.#1}
\newcommand{\wrt}[1][\ ]{w.r.t.#1}
\newcommand{\resp}[1][\ ]{resp.#1}

@@

Finally, we assume the existence of a universal data domain
given as a partially-ordered set $(\data, \order)$, potentially
encompassing several copies of any given data type with
different orders.  For example, we assume that $(\data,
\order)$ comprises both the unordered set of Booleans $\sB =
\tuple{\{\true, \false\}, \emptyset}$ and the naturally ordered one
$\sB^< = \tuple{\{\true, \false\}, \{\false < \true\}}$, and
similarly for integer and real numbers; as well as the set of
intervals ordered by inclusion.
%
\addSB{In practice, and in the examples of
this paper, all the variables are sorted, meaning that their values
can only range over given subsets of $\data$, such as $\sB$ or
$\sB^<$.
%
When speaking of an ordered sort, \eg $\sB^<$, we will assume that it
forms a meet-semilattice and denote by $\wedge$ the corresponding meet
operator.}

For a set of variables $V \subseteq \variables$, we denote
$\valuations{V} \bydef{=} \{\val{}{}: V \rightarrow \data\}$
the set of \emph{valuations} of the variables in $V$.
Valuations extend canonically to expressions.  Thus
$\val{}{}(x)$ and $\val{}{}(e)$ denote, respectively, the
values of the variable $x$ and the expression $e$ under the
valuation $\val{}{}$.
%
\addSB{Since we consider variables to be sorted, valuations
  applied to expressions are interpreted as partial functions
  $\exprs{V} \rightharpoonup \data$.}

For a valuation $\val{}{} \in
\valuations{V}$ and an assignment $(x_i := e_i)^{i \in I} \in
\assigns{V}$, we denote $\val{}{}[(x_i := e_i)^{i \in I}]$ the
valuation defined by
\[
\val{}{}\bigl[(x_i := e_i)^{i \in I}\bigr](x) \bydef{=}
\begin{cases}
  \val{}{}(x), & \text{if } x \not\in x_i^{i \in I}\,,\\
  \val{}{}(e_i), & \text{if } x = x_i, \text{for some } i \in I\,.
\end{cases}
\]
For two valuations $\val{1}{}, \val{2}{} : V \rightarrow
\data$, we denote $\valdiff{\val{1}{}}{\val{2}{}} \bydef{=}
\bsetdef{x \in V}{\val{1}{}(x) \neq \val{2}{}(x)}$ the set of
variables that are assigned different values by the two
valuations.  As usual, we write $\val{1}{} \order \val{2}{}$
iff $\val{1}{}(x) \order \val{2}{}(x)$, for all $x \in V$.  An
expression $e \in \exprs{V}$ is called \emph{monotonic} if, for
any two valuations $\val{1}{}, \val{2}{} \in \valuations{V}$,
$\val{1}{} \order \val{2}{}$ implies $\val{1}{}(e) \order
\val{2}{}(e)$.
%
Similarly, an assignment $(x_i := e_i)^{i \in I} \in \assigns{V}$ is
called monotonic if all expressions $e_i^{i \in I}$ are monotonic.
We will denote $\guards{V} \subset \boolexprs{V}$, $\monexprs{V}
\subset \exprs{V}$ and $\updates{V} \subset \assigns{V}$ the sets of
monotonic Boolean and generic expressions and assignments,
respectively.
%

@@

%% \noteSBin{%
%% To guarantee preservation of properties guards and expressions
%% have to be \emph{monotonic}.  In the case of a trivial order, \ie
%% when nothing is comparable, \emph{any guard or expression is
%%   monotonic}.  Thus, this is not really a constraint!

%% Do we put this in the definition of expressions and guards or in
%% the definitions of components and interaction models?
%% }

\begin{definition}[Components]
  \label{defn:component}
  A \emph{component} is a tuple $\tuple{Q, q^0, V, \val{0}{}, P,
  \export, \goesto{}}$, where
  \begin{itemize}
  \item $Q$ is a set of \emph{states}, with $q^0 \in Q$ the
    \emph{initial state}, 
  \item $V$ is a set of \emph{component variables},
  \item $\val{0}{} : V \rightarrow \data$ is an \emph{initial
    valuation} of the component variables, 
  \item $P$ is a set of \emph{ports}, with
    $\export : P \rightarrow 2^V$ 
    the \emph{set of variables exported by each port},
  \item $\goesto{}\, \subseteq
    Q \times (2^P\setminus \{\emptyset\}) \times \guards{V} \times \updates{V} \times Q$
%
    is a \emph{transition relation}, with transitions
    labelled by \emph{interactions}, \ie triples consisting of a non-empty set of ports,
    %% $\emptyset \neq a \subseteq P$,
    a monotonic Boolean \emph{guard}
    %% $g \in \guards{V}$ 
    and a monotonic \emph{update assignment}.
    %% \hl{$u \in \updates{V}$}.
  \end{itemize}
%
  We call the triple $(V,P, \export)$ the \emph{interface} of the
  component.\,\footnote{%
%
  In practice, only exported variables, \ie those in the set
  $\bigcup_{p \in P}\export(p)$, belong to the component interface and
  can be used for the interactions with other components (see
  \defn{im}).  However, in the context of this paper we can simplify
  by omitting this separation.
%
  }
%
  Notations $q \goesto{a, g, u} q'$ and $q \goesto{a, g, u}$ are as usual;
  for a component $B$, we denote $Q_B$, $q^0_B$, $V_B$, $\val{0}{B}$, $P_B$,
  and $\export[B]$ the corresponding constituents of $B$.  We will
  skip the index on the transition relations $\goesto{}$, since it
  is always clear from the context.
\end{definition}

\begin{figure}[t]
  \centering
  \includegraphics[width=0.9\columnwidth]{BIPspec-ArchFailureTimerMax-v4}
  \caption{The BIP specification of the Failure Monitor architecture}
  \label{fig:schema:ArchFailure:BIP}
\end{figure}

@@

The model in \fig{schema:ArchFailure:BIP} has  components
%% The
%% component $\NameOprnd$ is an \emph{operand} of the architecture, \ie
%% it only has its interface defined: $(\emptyset, \{\PortFinish,
%% \PortResume, \PortFail\}, \export[\emptyset])$, with
%% $\export[\emptyset](p) = \emptyset$, for all $p \in \{\PortFinish,
%% \PortResume, \PortFail\}$.  
%% The other two components,
%
$\NameTimer$(imer) and $\NameCtrl$(ontrol), with interfaces
$\btuple{
  \{t, z\},
  \{\PortTick, \PortCancel, \PortTOT\},
  \bigl\{\PortTick \mapsto \{t, z\}\bigr\}
}$,
and
$\btuple{
  \{\mathit{zone}\},
  \{\PortReset, \PortFail, \PortResume, \PortTOC\},
  \bigl\{\PortFail \mapsto \{\mathit{zone}\}\bigr\}
}$, respectively.
%
%% \[
%% \begin{array}{@{}c@{\hspace{3mm}}c@{}}
%%   \bigl(
%%   \{t, z\},
%%   \{\PortTick, \PortCancel, \PortTOT\},
%%   \export[\NameTimer]
%%   \bigr)\,,
%%   %% \qquad
%%   &
%%   \bigl(
%%   \{\mathit{zone}\},
%%   \{\PortReset, \PortFail, \PortResume, \PortTOC\},
%%   \export[\NameCtrl]
%%   \bigr)\,,
%%   \\
%%   \export[\NameTimer](p) =
%%   \begin{cases}
%%     \{t, z\}, & \text{if } p = \PortTick,\\
%%     \emptyset, & \text{otherwise,}
%%   \end{cases}
%%   %% \]
%%   %% and
%%   %% \[
%%   %% \quad
%%   &
%%   \export[\NameCtrl](p) =
%%   \begin{cases}
%%     \{\mathit{zone}\}, & \text{if } p = \PortFail,\\
%%     \emptyset, & \text{otherwise.}
%%   \end{cases}
%% \end{array}
%% \]
%
%% , where $\export[\NameCtrl](\PortFail) =
%% \{\mathit{zone}\}$ and $\export[\NameCtrl](p) = \emptyset$, for all
%% other ports $p$; $\export[\NameTimer](\PortTick) = \{t, z\}$ and
%% $\export[\NameTimer](p) = \emptyset$, for all other ports $p$.
Variable $t$ is implicitly assumed to be an integer.  Variables $z$
and $\mathit{zone}$ are implicitly assumed to be integer intervals of
the form $[l,u]$.  For both types, we also implicitly assume the
natural ordering: increasing integers and interval inclusion.

%% are \emph{coordinators}\mdash they are specified
%% completely, including the states and transitions defining their
%% \emph{behaviour}.
Component \emph{behaviour} is defined by states %% ($\{\cstate{0},
%% \cstate{1}, \cstate{2}\}$, for $\NameCtrl$; $\{\tstate{0}\}$, for
%% $\NameTimer$)
and transitions.
The initial states
\tstate{0} and \cstate{0}, and valuations $\val{0}{\NameTimer} = \{t
\mapsto 0, z \mapsto \topinterval\}$ and $\val{0}{\NameCtrl} =
\{\mathit{zone} \mapsto [\mathrm{Min}, \mathrm{Max}]\}$ are shown by
the incoming arrows $\goesto{\TimerInit[]} \tstate{0}$ and
$\goesto{\IntervalInit[]} \cstate{0}$.\footnote{%
%
  We use $\topinterval$ to denote the top interval in the inclusion
  order, \ie $(-\infty, +\infty)$.
%
} The constants $\mathrm{Min}$ and $\mathrm{Max}$ are the parameters
of the architecture.

Transitions are labelled with ports of the
corresponding components, Boolean guards and update assignments on local
variables.  \Eg[,] the loop transition $\tstate{0}
\goesto{\PortTick, \GuardTick[], \TimerUpd[]} \tstate{0}$ in the
component $\NameTimer$ is labeled by the port $\PortTick$.  It can be
fired only when the current value of the local variable $t$ is less
than $z.u$.
%% , where $z = [l, u]$ is a local variable representing an
%% interval
Upon firing, this transition increments the value of $t$ by $1$.  The
guards and update assignments of the transitions of $\NameCtrl$ are
omitted.  The default values are the constant predicate $\true$ (true) and
the empty assignment {\noop}, respectively.  
%\noteLH{It looks like Max is in zone, I do not understand}
%
Clearly, all guards and update assignments are monotonic.

%% \todoSBin{Speak of guard monotonicity}

\begin{definition}[Component semantics]
  \label{defn:comp:semantics}
  The \emph{semantics} of a component $B = \tuple{Q, q^0, V,
  \val{0}{}, P, \export, \goesto{}}$ is %% given by
  the LTS denoted $\semopen{B} = \tuple{S,
  s^0, \goesto{}}$, where $S = Q \times \valuations{V}$, $s^0 =
  \tuple{q^0, \val{0}{}}$ and $\goesto{}$ is the minimal transition
  relation satisfying the rule
  %
  \begin{equation}
    \label{eq:comp:semantics}
    \infer{
      q \goesto{a, g, u} q'
      \and
      \val{}{} \models g
      \and
      \val[\primeit]{}{} = \tilde{\val{}{}}[u]
      \and
      \valdiff{\val{}{}}{\tilde{\val{}{}}} \subseteq \export(a)
    }{
      (q, \val{}{}) \goesto{a,\tilde{\val{}{}}} (q', \val[\primeit]{}{})
    }
    \,.
  \end{equation}
  %
  %% \todoSB{Do we actually need this?}\todoLH{I agree: I do not think we need this}
  %% \hl{The \emph{closed semantics} of $B$ is given by the LTS denoted
  %% $\semclosed{B} \bydef{=}\reachable{\semopen{B}}$, comprising
  %% only the reachable states of $\semopen{B}$.}
\end{definition}

@@

\begin{definition}[Interaction model]
  \label{defn:im}
  For a finite set of component interfaces $(V_i, P_i,
  \export[i])^{i \in I}$, such that all $P_i$ and all $V_i$ are
  pairwise disjoint, %% \ie $\forall i \neq j,\ P_i \cap P_j = V_i \cap
  %% V_j = \emptyset$. 
  let $P = \bigcup_{i \in I} P_i$, $V = \bigcup_{i
    \in I} V_i$ and $\export : P \rightarrow 2^V$ such that, for any
  $p \in P_i$, $\export(p) = \export[i](p)$. 
%
  %% \noteSB{Updated this}\hl{and $V_a \bydef{=} \bigcup_{i \in
  %%   \supp{a}} \export[i](a \cap P_i)$ the set of component
  %% variables exported through $a$.}

@@

\begin{definition}[Composition]
  \label{defn:im:sem}
  The \emph{composition} of a finite set of components $\cB = \tuple{Q_i,
  q^0_i, V_i, \val{0}{i}, P_i, \export[i], \goesto{}}^{i \in I}$ with
  the interaction model $\Gamma$ over $\tuple{V, P, \export}$ (as in \defn{im})
  is the component $\Gamma(\cB) = \tuple{Q, q^0, V, \val{0}{}, P,
    \export, \goesto{}}$, where
%
  $Q = \prod_{i \in I} Q_i$;
%
  $q^0 = (q_i^0)^{i \in I}$;
%
  $\val{0}{}: V \rightarrow \data$ is such that, for any $v \in
  V_i$, $\val{0}{}(v) = \val{0}{i}(v)$;
%
  and $\goesto{}$ is the minimal transition relation satisfying
  the rule
%
  \begin{gather*}
%%     \label{eq:im:empty}
%%     \infer{
%%       q_j \goesto{\emptyset, g, e} q_j'
%%       \and
%%       \forall i \neq j, q_i = q_i'
%%     }{
%%       (q_i)^{i \in I} \goesto{\emptyset, g, e} (q_i')^{i \in I}
%%     }\,,
%% %
%%     \\[0.5\baselineskip]
    \label{eq:im:sem}
%
    \infer{    
      (a, g, u) \in \Gamma
      \and
      a \neq \emptyset
      \and
      \forall i \in \supp{a}, q_i \goesto{a \cap P_i, g_i, u_i} q_i'
      \and
      \forall i \not\in \supp{a}, q_i = q_i'
%
      \\\\
      \textstyle
%
      g' = g \land \bigwedge_{i \in \supp{a}} g_i
      \and
      u' = u; u_i^{i \in \supp{a}}
    }{
      (q_i)^{i \in I} \goesto{a, g', u'} (q_i')^{i \in I}
    }\,.
  \end{gather*}
  %% \todoSBin{Add the intuitive explanation}
\end{definition}

\addSB{Intuitively, an interaction can be fired if all the involved
  components are ready to fire their corresponding transitions.  The
  other components do not change their states.  Both the interaction
  guard and those of the participating transitions must be satisfied.
  The update assignment of the interaction is executed first, followed
  by those of the components.}

@@

Connectors are hierarchical, tree-like structures with component ports
at the leaves.  They define sets of interactions, based on the
attributes of the nodes, which may be either \emph{trigger}
(triangles in \fig{schema:ArchFailure:BIP}) or \emph{synchron}
(bullets in \fig{schema:ArchFailure:BIP}).
%
If all sub-connectors of a connector are synchrons, then an
interaction is allowed by the connector only if each subconnector
can contribute.
%
If at least one of the sub-connectors is a trigger, then any
interaction consisting of contributions of any set of sub-connectors
\emph{involving at least one of the triggers} is allowed.
%
The interaction model is defined as the set of all interactions
allowed by at least one of the connectors in the system.
%% The composition semantics of BIP systems consists in
%% firing exactly one interaction, enabled through at least one of the
%% top-level connectors, at each execution round.

For instance, the connector {\NameTimer.\PortTick
  \trigsynch (\PortFail \trigsynch \NameCtrl.\PortFail)} is a
two-level hierarchical connector.  In the second-level subconnector
{\PortFail \trigsynch \NameCtrl.\PortFail}, the port {\PortFail} is a
trigger, whereas {\NameCtrl.\PortFail} is a synchron.  Therefore, this
subconnector allows two interactions labelled $\{\PortFail\}$ and $\listset{\NameCtrl.\PortFail, \PortFail}$.  Similarly, at the top level,
{\NameTimer.\PortTick} is a trigger, whereas the second-level
subconnector is a synchron.  Hence, the overall connector allows the
interaction involving only one port {\NameTimer.\PortTick}\footnote{%
%
  Such interactions are called \emph{singleton interactions}.
%
} and any interaction involving both {\NameTimer.\PortTick} and
\emph{some} interaction of the second-level subconnector.  Combining
the two levels we can see that the entire connector defines the
following three interactions (observe that $\topinterval +
\NameTimer.t = \topinterval$ and $\NameTimer.z \cap \topinterval =
\NameTimer.z$):
%
\begin{gather*}
  \bigl(\{\NameTimer.\PortTick\},
  \true, \emptyset\bigr)
  %% \NameTimer.z := \NameTimer.z\bigr)
  ,
%
  \qquad
%
  \bigl(\{\PortFail, \NameTimer.\PortTick\},
  \true, \emptyset\bigr)
  %% \NameTimer.z := \NameTimer.z\bigr)
  ,
%
  \\
%
  \bigl(\{\NameCtrl.\PortFail, \PortFail, \NameTimer.\PortTick \},
  \true,
  \NameTimer.z := \NameTimer.z \cap (\NameCtrl.\mathit{zone} + \NameTimer.t)\bigr)
  .
\end{gather*}
%% \begin{align*}
%%   \bigl(\{\NameTimer.\PortTick\}&,
%%   \true, 
%%   \NameTimer.z := \NameTimer.z\bigr)
%%   \,,\\
%%   \bigl(\{\PortFail, \NameTimer.\PortTick\}&,
%%   \true,
%%   \NameTimer.z := \NameTimer.z\bigr)
%%   \,,\\
%%   \bigl(\{\NameCtrl.\PortFail, \PortFail, \NameTimer.\PortTick \}&,
%%   \true,
%%   \NameTimer.z := \NameTimer.z \cap (\NameCtrl.\mathit{zone} + \NameTimer.t)\bigr)
%%   \,.
%% \end{align*}

%% \todoSB{Update this}
%% \hl{For instance, in \fig{schema:ArchFailure:BIP}, the two ports
%% $\NameTimer.\PortStart$ and $\NameCtrl.\PortFail$ are always
%% synchronised, since they belong to exactly one binary sub-connector,
%% where they are both synchrons.  In particular, this means that
%% whenever the transition \cstate{0} \goesto{\PortFail} \cstate{1} is
%% fired, so is the transition \tstate{0} \goesto{\PortStart,
%%   \TimerInit[]} \tstate{1}, initialising the timer.  The binary
%% connector $\NameTimer.\PortStart \twosynch
%% \NameCtrl.\PortFail$ is a sub-connector of a hierarchical connector,
%% where the port $\NameOprnd.\PortFail$ is a trigger.  Thus, the above
%% interaction can only happen together with $\NameOprnd.\PortFail$,
%% forming a ternary interaction.  On the contrary, being a trigger, the
%% port $\NameOprnd.\PortFail$ can fire alone, forming a singleton
%% interaction.}

%% \todoSBin{Speak of data transfer}

%% \todoSBin{Provide the interaction model}

@@

%% \begin{note}
%%   \label{rem:maxprogress}
%% \end{note}

@@

Architectures are partially defined BIP models, where \emph{dangling}
ports serve as placeholders for the eventual connection with
additional \emph{operand} components.
